# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?

* First filter the data with no NAs then calculate the total number of steps taken per day.
```{r}
filtered <- activity[!is.na(activity$steps), ]
head(filtered)
summary(filtered)
str(filtered$date)
library(lubridate)
filtered$date <- ymd(filtered$date)
str(filtered$date)
steps.everyday <- sapply(split(filtered$steps, filtered$date), sum)
uni.dates <- unique(filtered$date)
steps.per.day <- data.frame(uni.dates, steps.everyday)
names(steps.per.day) <- c("date", "steps")
head(steps.per.day)
```

* Plot steps taken per day
```{r, fig.width = 9}
library(lattice)
xyplot(steps ~ date, data = steps.per.day, type = "l", main = "Steps Taken Each Day", scales = list(tick.number = 10))
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
