# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?


* First filter the data with no NAs then calculate the total number of steps taken per day.
```{r}
filtered <- activity[!is.na(activity$steps), ]
head(filtered)
summary(filtered)
str(filtered$date)
library(lubridate)
filtered$date <- ymd(filtered$date)
str(filtered$date)
steps.everyday <- sapply(split(filtered$steps, filtered$date), sum)
uni.dates <- unique(filtered$date)
steps.per.day <- data.frame(uni.dates, steps.everyday)
names(steps.per.day) <- c("date", "steps")
head(steps.per.day)
```


* Plot steps taken per day
```{r, fig.width = 7, fig.height = 4, fig.align='center'}
hist(steps.per.day$steps, breaks = seq(0, 25000, 2500), xaxt = 'n', 
     main = "Histogram of steps taken each day")
axis(side = 1, at = seq(0, 25000, 2500))
```
```{r, fig.width = 7, fig.height = 4, fig.align='center'}
library(ggplot2)
qplot(x = date, y = steps, data = steps.per.day, geom = "line")
```


* Calculate the mean and median total number of steps taken per day
```{r}
print(paste("the mean is:", mean(steps.everyday)))
print(paste("the median is:", median(steps.everyday)))
```


## What is the average daily activity pattern?

* Generate two corresponding daily pattern data frames concerning the mean and median of steps taken with respect of the time interval identifier
```{r}
library(reshape2)
act.melt <- melt(activity, id = c("interval", "date"), 
                 measure.vars = "steps")
pattern.mean <- dcast(act.melt, interval ~ variable, mean, na.rm = T)
pattern.median <- dcast(act.melt, interval ~ variable, median, na.rm = T)
head(pattern.mean)
head(pattern.median)
```


* Plot the two graphs
```{r, fig.width = 7, fig.height = 4, fig.align='center'}
ggplot(aes(x = interval, y = steps), data = pattern.mean) +
    geom_line(color = "firebrick3")
ggplot(aes(x = interval, y = steps), data = pattern.median) +
    geom_line(color = "dodgerblue2")
```


* Infer which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.
```{r}
ordered.pattern <- pattern.mean[order(pattern.mean$steps, decreasing = T), ]
head(ordered.pattern)
```
```{r, echo = FALSE}
max.interval <- ordered.pattern$interval[1]
```

*Interval `r max.interval` contains the maximum number of steps*


## Imputing missing values
```{r}
na.number <- table(complete.cases(activity))[["FALSE"]]
```
*the total number of missing values in the dataset is `r na.number`*

I'm using the mean for the particular 5-minute interval to fill in the missing value
```{r}
activity.full <- activity
activity.full$date <- ymd(activity.full$date)
indice <- which(is.na(activity.full$steps))
for(i in 1:length(activity.full[indice, ]$steps)){
    activity.full[indice, ]$steps[i] <- 
        pattern.mean[pattern.mean$interval == 
                         activity.full[indice, ]$interval[i], ]$steps
}
```
* Histogram of the total number of steps taken each day
```{r, fig.width = 7, fig.height = 4, fig.align='center'}
full.steps.everyday <- sapply(split(activity.full$steps, 
                                    activity.full$date), sum)
range(full.steps.everyday)
hist(full.steps.everyday, breaks = seq(0, 25000, 2500), xaxt="n", 
     main = "Histogram of steps taken per day with data filled in")
axis(1, at = seq(0, 25000, 2500))
```
* Calculate the mean and median of steps taken per day
```{r}
ans1 <- as.character(round(mean(full.steps.everyday)))
ans2 <- as.character(round(median(full.steps.everyday)))
```
*The mean is `r ans1`. The median is `r ans2`*
*The relative distribution changes and the overall steps taken increases*

## Are there differences in activity patterns between weekdays and weekends?
* Melt and cast the data then plot
```{r, fig.width = 9, fig.height = 4, fig.align='center'}
activity.full$dob <- tolower(weekdays(activity.full$date))
for(i in 1:nrow(activity)){
    if(activity.full$dob[i] == "saturday" | 
           activity.full$dob[i] == "sunday"){
        activity.full$week[i] <- "weekend"
    }else{
        activity.full$week[i] <- "weekday"
    }
}
activity.full$week <- factor(activity.full$week)

full.melt <- melt(activity.full, id = c("interval", "date", "week"), 
                 measure.vars = "steps")
new.pattern <- dcast(full.melt, interval + week ~ variable, mean)

ggplot(aes(x = interval, y = steps), data = new.pattern) +
    geom_line(aes(color = week)) + facet_wrap( ~week) +
    theme(legend.title=element_blank(),  axis.text=element_text(size=12),
        axis.title=element_text(size=14))
```

